### 2020-03-24 response.txt

We would like to thank the anonymous reviewers for their helpful comments.

Below are responses to questions and clarifications.

R1:

Most importantly: https://github.com/mutantsforstaticanalysis/data now provides all the raw data used in the paper.  We hope this helps with qualitiative review.  We also, below, discuss briefly the issue of whether the mutants detected by any tools were of actual interest, which is indeed a core issue.  We do not examine mutants detected by no tools; due to the differential nature of the approach, we consider these potentially noise/uninteresting/not plausibly statically detected, but can simply ignore them.  Only when at least one tool detects a mutant do we care.  We note that the set of mutation operators can be easily extracted by examining the rule files in the open source repository for universalmutator.

Yes, we actually used "number" of warnings, as a conservative measure of detection.  If a mutant simply causes a tool to change the message it reports, then we do not consider that a kill.  Similarly, if it loses one warning and produces another (which is in a sense the same thing, from a user's perspective), then it is not a kill.  The core reason we don't consider the location/nature of the warning change is that in some cases a mutant causes a warning about code far from the mutant location.  Therefore, we strictly require that a tool produce _more warnings_ when code is mutated, in order to be counted as detecting the mutant.  We experimented with other definitions on a small set of smart contracts, and this was much better than other approaches, in terms of the results making sense.  It is possible (though we didn't see this in practice with smart contracts) that it undercounts cases where 1) the original code produces warning A and 2) the mutated code "fixes" problem A and causes detection of a different problem, B.  If the mutant doesn't change the issue, just changes the exact nature of the problem, then we argue the mutant didn't introduce a fault, it just changed some aspect of an existing fault, which is not what we are interested in.  If we want to compare tools over existing faults, that is already doable.

Good point, the source/bytecode choice is language dependent.  As to why we chose slither, see below.

R2:

We chose PMD primarily because we started by focusing on [43] as a comparison paper, rather than [28], due to the highly-focused scope of [28] (only Defects4J bugs, thus many many fewer relevant warnings/aspects of code considered, and any biases in producing the Defects4J data set).  Adding Error Prone as well would be a good idea.  In a way, [43] is a closer comparison to our work:  larger scope of interest, if higher risk of invalid results over that larger scope.

"Importing" rules from one tool to another can be very helpful for actual users.  If one tool uses a better underlying analysis engine (e.g., better intermediate language, better alias/data-flow/etc. approach), then adding a rule from another tool can result in much better _actual_ detection of bugs, or fewer false positives.  Slither's IR and analysis are probably more scalable than Securify and more precise than SmartCheck, so even just using the same rule gives everyone better analysis results.  Also, the rules are generalizations that expand on the rules in SmartCheck/Securify: the mutation approach makes generalizing a specific rule somewhat easier, we think, due to providing multiple examples, not just a high-level concept.

We used universalmutator for two reasons:  primarily because PIT is a bytecode-level mutation tool only, PMD requires source code, and we believe users want to inspect mutants for this application (and we needed to manually do some validation that the differences were meaningful).  Second because we wanted to use a consistent set of operators and mutation approach, and universalmutator was the best tool for the other languages.

"better" in RQ4 means that we hypothesized that tools would produce _higher_ mutation scores (kill more mutants) for programs for which there were no findings in the original code.

Securify is a static analysis tool, but it "symbolically analyzes the contract’s dependency graph to extract
precise semantic information from the code" before checking compliance/violation patterns.  The first step, unfortunately, like many symbolic approaches, can sometimes scale abominably on real code.

Our reasoning on the mutant ratio for Securify is that Securify does not notably produce warnings for mutants (which are probably bugs in some sense, even if not statically detectable) compared to the vast number of warnings it issues in general.  Other research supports the claim that Securify is prone to false positives over all code, and its behavior on mutants is not appreciably different (see the 20x or more higher ratios for the other tools.

R3:

Yes, we had to understand each tool's types of messages (for Pylint this was quite a painful process) and filter out ones that are essentially coding style rather than "this could be a bug!" issues, to the extent possible.  We made considerably effort to ensure these choices were fair to the tools.  Some tools made it easy by simply providing a setting to turn off "informational and style" issues, but for others it is a (minor in our opinion) threat to validity.

### 2020-03-25 response.txt

We would like to thank the anonymous reviewers for their helpful comments.

Below are responses to questions and clarifications.

R1:

Most importantly: https://github.com/mutantsforstaticanalysis/data now provides all the raw data used in the paper.  We hope this helps with qualitiative review.  We also, below, discuss briefly the issue of whether the mutants detected by any tools were of actual interest, which is indeed a core issue.  We do not examine mutants detected by no tools; due to the differential nature of the approach, we consider these potentially noise/uninteresting/not plausibly statically detected, but can simply ignore them.  Only when at least one tool detects a mutant do we care.  The full set of mutation operators can be easily extracted by examining the rule files in universalmutator github repo.

Qualitative analysis is hard to make completely scientific, but we agree the paper omitted some of the forest for the trees, to make way for details of comparisons.  We did in fact inspect a large number of the mutants in the "diff" (via random sampling for Java/Python).  For Solidity, where we manually examined the majority of the mutants killed by at least one tool, most appeared to be semantically "troubling" at minimum, or clearly faults, though we think many changes of msg.sender to tx.origin flagged by non-Slither tools may be valid (if strange) code.  For Java and Python, many mutants detected by one tool but not another involved deleted function calls or conditional changes (e.g. == null to != null) that would clearly introduce potential null pointer exceptions or method calls on None.  These show differences in analysis methods rather than missing rules, but are clearly real(istic) faults.  Other Java/Python cases were harder to understand, but many seemed to be real faults (e.g., introducing possible division by zero).

### 2020-03-25 response.txt

We would like to thank the anonymous reviewers for their helpful comments.

R1:

Most importantly: https://github.com/mutantsforstaticanalysis/data has the raw data.  We hope this helps!  The full set of mutation operators can be easily extracted by examining the rule files in universalmutator github repo.

With respect to the meaningfulness of the mutants, first note that we do not examine mutants detected by zero tools, due to the differential nature of the approach; these are likely noise/uninteresting/not plausibly statically detected, but we just ignore them.  Only when at least one tool detects a mutant do we care.

Qualitative analysis is hard to make completely scientific, but we agree the paper omitted the forest for the trees, to make way for details of tool comparisons.  We did in fact inspect a large number of the mutants in the "diff" (via random sampling for Java/Python).  For Solidity, where we manually examined the majority of the mutants killed by at least one tool, _most appeared to be semantically "troubling" at minimum_, or clearly faults, though we think many changes of msg.sender to tx.origin flagged by non-Slither tools may be valid (if strange) code.  For Java and Python, many mutants detected by one tool but not another involved deleted function calls or conditional changes (e.g. == null to != null) that would clearly introduce potential null pointer exceptions or method calls on None.  These show differences in analysis methods rather than missing rules, but are clearly real(istic) faults.  Other Java/Python cases were harder to understand, but many seemed to be real, statically detectable faults (e.g., introducing possible division by zero).

@@

Yes, we actually used "number" of warnings, as a conservative measure of detection.  If a mutant simply causes a tool to change the message it reports, then we do not consider that a kill.  Similarly, if it loses one warning and produces another (which is in a sense the same thing, from a user's perspective), then it is not a kill.  The core reason we don't consider the location/nature of the warning change is that in some cases a mutant causes a warning about code far from the mutant location.  Therefore, we strictly require that a tool produce _more warnings_ when code is mutated, in order to be counted as detecting the mutant.  We experimented with other definitions on a small set of smart contracts, and this was much better than other approaches, in terms of the results making sense.  It is possible (though we didn't see this in practice with smart contracts) that it undercounts cases where 1) the original code produces warning A and 2) the mutated code "fixes" problem A and causes detection of a different problem, B.  If the mutant doesn't change the issue, just changes the exact nature of the problem, then we argue the mutant didn't introduce a fault, it just changed some aspect of an existing fault, which is not what we are interested in.  If we want to compare tools over existing faults, that is already doable.

Good point, the source/bytecode choice is language dependent.  As to why we chose slither, see below.

R2:

We chose PMD primarily because we started by focusing on [43] as a comparison paper, rather than [28], due to the highly-focused scope of [28] (only Defects4J bugs, thus many many fewer relevant warnings/aspects of code considered, and any biases in producing the Defects4J data set).  Adding Error Prone as well would be a good idea.  In a way, [43] is a closer comparison to our work:  larger scope of interest, if higher risk of invalid results over that scope.

@@

"Importing" rules from one tool to another can be very helpful for actual users.  If one tool uses a better underlying analysis engine (e.g., better intermediate language, better alias/data-flow/etc. approach), then adding a rule from another tool can result in much better _actual_ detection of bugs, or fewer false positives.  Slither's IR and analysis are probably more scalable than Securify and more precise than SmartCheck, so even just using the same rule gives everyone better analysis results.  Also, the rules are generalizations that expand on the rules in SmartCheck/Securify: the mutation approach makes generalizing a specific rule somewhat easier, we think, due to providing multiple examples, not just a high-level concept.

We used universalmutator for two reasons:  primarily because PIT is a bytecode-level mutation tool only, PMD requires source code, and we believe users want to inspect mutants for this application (and we needed to manually do validation as R1 points out, which is hard for bytecode).  We also wanted to use a consistent set of operators and mutation approach, and universalmutator was the best tool for the other languages.

In RQ4 "better" means _higher_ mutation scores were hypothesized for programs for which there were no findings in the original code.

Securify is a static analysis tool, but it "symbolically analyzes the contract’s dependency graph to extract precise semantic information from the code" before checking compliance/violation patterns.  The first step, unfortunately, can sometimes scale abominably on real code.

@@

Our reasoning on the mutant ratio for Securify is that Securify does not notably produce warnings for mutants (which are probably bugs in some sense, even if not statically detectable) compared to the vast number of warnings it issues in general.  Other research supports the claim that Securify is prone to false positives over all code, and its behavior on mutants is not appreciably different (see the 20x or more higher ratios for the other tools.

R3:

Yes, we had to understand each tool's finding types (for Pylint this was quite painful) and filter out ones that are essentially stylistic rather than "here’s a bug!".  We made considerable effort to ensure these choices were fair to the tools.  Some tools made it easy by simply providing a setting to turn off "informational and style" issues, but for others it is a (minor, we think) threat to validity.

### 2020-03-27 response.txt

We thank the reviewers for their helpful comments.

R1:

Most importantly: https://github.com/mutantsforstaticanalysis/rawdata has the raw data. We hope this helps! The full set of mutation operators can be easily extracted by examining the rule files in the universalmutator github repo.

With respect to the meaningfulness of the mutants, first note that we do not examine mutants detected by zero tools, due to the differential nature of the approach; these are likely noise/uninteresting/not plausibly statically detected, but we just ignore them. Only when at least one tool detects a mutant do we care about it (and mostly when at least one does not).

Qualitative analysis is hard to make completely scientific, but we agree the paper omitted the forest for the trees, to make way for details of tool comparisons. We did in fact inspect a large number of the mutants in the "diff" (via random sampling for Java/Python). For Solidity, where we manually examined the majority of the mutants killed by at least one tool, _most appeared to be semantically "troubling" at minimum_, or clearly faults, though we think many changes of msg.sender to tx.origin flagged by non-Slither tools may be valid (if strange) code. For Java, a large number of mutants detected by one tool but not another involved deleted function calls or conditional changes (e.g. == null to != null) that introduce potential null pointer exceptions. These show differences in analysis methods rather than missing detectors, and are clearly faulty code that should be detected. For Python, mutants producing code that would cause runtime (but not parsing) failures were common, often just an access to a now-undefined variable or obvious type mismatch. Again, all tools aim to detect these, so it is engine more than detector differences. Other Java/Python cases were harder to understand, but many seemed to be real, statically detectable faults. We will add this type of qualitative analysis.

Yes, we actually used "number" of warnings, as a conservative measure of detection. If a mutant only causes a tool to change finding message, then we do not consider that a kill. If it loses one warning and produces another then it is not a kill. The core reason we don't consider the location/nature of the warning change is that in some cases a mutant should cause a warning about code far from the mutant location. Therefore, we strictly require that a tool produce _more warnings_ when code is mutated, in order to be counted as detecting the mutant; we assume each mutant of interest potentially adds a fault, and fixes no existing problems. We experimented with other definitions on a set of smart contracts, and this approach produced the most useful results. It is thus possible (though we didn't see this happen with smart contracts) that we undercount cases where the original code produces warning A and the mutated code "fixes" problem A but introduces problem B. This seems like an acceptable risk to take, given mutants seldom fix faults, and better than having to parse/understand warning types/content to apply the method to every new language/tool.

Good point, the source/bytecode choice is language dependent. As to why we chose Slither, see below.

We do observe project level differences, and will report them.

While not for Java, universalmutator has been extensively tuned for Python.

We’ll add those references!

@@

R2:

We chose PMD primarily because we started by focusing on [43] as a comparison paper, rather than [28], due to the highly-focused scope of [28] (only Defects4J bugs, thus many many fewer relevant warnings/aspects of code considered, and any biases in producing the Defects4J data set). Adding Error Prone would be good, and should be possible by camera-ready. Arguably, [43] is the closer comparison to our work: larger scope of interest, albeit with higher risk of invalid results over that scope.

"Importing" rules from one tool to another can be very helpful for users. If one tool uses a better underlying analysis engine (e.g., better intermediate language, better alias/data-flow/etc. approach), then adding a rule from another tool can result in much better _actual_ detection of bugs, or fewer false positives. Slither's IR and analysis are probably more scalable than Securify and more precise than SmartCheck, so even just using the same rules gives everyone better analysis results. Also, the rules are generalizations that expand on the rules in SmartCheck/Securify: using mutations makes generalizing a specific rule somewhat easier, we believe, due to (often multiple) concrete examples, not just a high-level concept.

We primarily used universalmutator because PIT is a bytecode-level mutation tool only. PMD requires source code, and we believe users want to inspect some mutants. We needed to manually do validation, as R1 points out, which is hard for bytecode. We wanted to use a consistent set of operators and mutation approach, and universalmutator was the best-supported tool for the other languages.

@@

In RQ4 "better" means _higher_ mutation scores were hypothesized for programs for which there were no findings in the original code.

Securify is a static analysis tool, but it uses souffle under the hood to solve Prolog constraints as part of building a dependency graph; this can scale very poorly.

Our reasoning on the mutant ratio for Securify is that Securify does not notably produce more warnings for mutants (which are usually bugs in some sense, even if not statically detectable) compared to the vast number of warnings it issues in general. Other research supports the claim that Securify is prone to false positives, and its behavior on mutants is not appreciably different than on other code, while SmartCheck and Slither have 20x better ratios.

@@

R3:

Yes, we had to understand each tool's finding types (for Pylint this was painful) and filter out ones that are essentially stylistic rather than "here’s a bug!". We made considerable effort to ensure these choices were fair to the tools. Some tools made it easy by simply providing a setting to turn off "informational and style" issues, but for others it is a (minor, we think) threat to validity.

### 2020-04-15 experiments.tex
current draft

\begin{itemize}
\item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?  That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar in terms of mutation detection capability?
\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?  Do \emph{ratios} differ between programs, or does the number of mutants killed simply reflect the ``verbosity'' of each tool?
\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?  Can we confirm results from other evaluation methods, or informal opinion?
\item {\bf RQ4:}  Are mutation scores for clean programs better than those for programs for which tools report findings before the program is mutated?
\item {\bf RQ5:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
 different tools are good at/bad at, and extend this information to improve tools by addressing identified weaknesses?
\item {\bf RQ6:}  How easy is it identify low-hanging fruit for improving tools \emph{without} prioritization to remove conceptually redundant mutants?
  \end{itemize}

In particular, we consider {\bf RQ2} to be of critical importance; if the mutant ratios for tools differ, then this is clear evidence that our hypothesis that the tendency of mutants to be faults, and to expect that mutated code will, by a more precise and accurate tool, be flagged as problematic more often than non-mutated code, holds.  This expectation that (some subset of the) mutants can serve as proxies for real, detectable faults is the core concept of our approach.
{\bf RQ4} addresses a concern briefly mentioned in the introduction:  it is possible that warnings for the original code interfere with our definition of detection.  The ideal case for our approach is when a tool report no findings for un-mutated code, and either reports a finding (kills the mutant) or not (does not kill it) when the mutant is introduced.  Chekam et al. showed that the ``clean program assumption'' for testing, the idea that coverage will be similar for faulty and fixed versions of a program, is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, and we want to establish that this is unlikely to be the case for our approach.
{\bf RQ5} and {\bf RQ6} receive only preliminary answers, for the smart contract tools.

@@

For {\bf RQ4}, on the changes seen when restricting analysis to clean contracts, Slither did slightly better at detecting mutants when the original contract was clean for Slither, and the other two tools did somewhat worse on contracts for which they reported no findings.  For the three contracts clean for all tools, Slither performed almost exactly as it did over contracts in general, and the other tools performed worse, by about the same margin as they did for their own clean contracts.  For our approach, we only need a week version of the ``clean program assumption'':  the threat is that kills may be under-reported for non-clean programs, due to interference with findings for the original code.  It is not a problem if mutation scores are \emph{worse} for programs where a tool reports no findings for the un-mutated code.  We therefore, for smart contracts, find no threat to our approach arising from the presence of findings on un-mutated code.  We speculate that ``clean'' results for some tools result from contracts where the tool has trouble with the contract code, but does not actually crash; Slither may do better on clean code because it has fewer such failures, and clean contracts are probably generally simpler and easier to analyze.

\subsubsection{Improving Slither {\bf (RQ5 and RQ6)}}

@@

Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants, suggesting that for {\bf RQ6} it is indeed hard to use mutation analysis results without prioritization. A large majority of the mutants we inspected involved either the missing
{\tt return} problem noted in the introduction, or replacing {\tt msg.sender} with {\tt tx.origin}; Slither, of course, has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of the differences with respect to such changes are due to intentional behavior. SmartCheck and Securify tend to identify most (though not all) uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule, intended to reduce false positives.
It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ5} and {\bf RQ6}.

@@

@@

@@

We used the Universal Mutator \cite{universalmutator,regexpMut}, which aggressively produces large numbers of mutants, but does not target any particular software defect patterns, to generate all mutants.  We will make available a repository containing our full raw results plus details on contracts and projects selected, for replication and further analysis, upon acceptance (to avoid breaking blinding).

>>>>

\begin{itemize}
\item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?  That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar in terms of mutation detection capability?
\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?  Do \emph{ratios} differ between programs, or does the number of mutants killed simply reflect the ``verbosity'' of each tool?
\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?  Can we confirm results from other evaluation methods, or informal opinion?
\item {\bf RQ4:}  Are mutation scores for clean programs higher (more mutants killed) than those for programs for which tools report findings before mutation?
\item {\bf RQ5:}  Are killed mutants reasonablely similar to actual faults?  That is, is the data above based on meaningful changes to code?
\item {\bf RQ6:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
 different tools are good at/bad at, and extend this information to improve tools by addressing identified weaknesses?
\item {\bf RQ7:}  How easy is it to identify low-hanging fruit for improving tools \emph{without} prioritization to remove conceptually redundant mutants?
  \end{itemize}

In particular, we consider {\bf RQ2} to be of critical importance; if the mutant ratios for tools differ, then this is clear evidence that our hypothesis that the tendency of mutants to be faults, and to expect that mutated code will, by a more precise and accurate tool, be flagged as problematic more often than non-mutated code, holds.  This expectation that (some subset of the) mutants can serve as proxies for real, detectable faults is the core concept of our approach.
{\bf RQ4} addresses a concern briefly mentioned in the introduction:  it is possible that warnings for the original code interfere with our definition of detection.  The ideal case for our approach is when a tool report no findings for un-mutated code, and either reports a finding (kills the mutant) or not (does not kill it) when the mutant is introduced.  Chekam et al. showed that the ``clean program assumption'' for testing, the idea that coverage will be similar for faulty and fixed versions of a program, is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, and we want to establish that this is unlikely to be the case for our approach.  {\bf RQ5} is somewhat inherently qualitative and incomplete; we cannot analyze all mutants on which results are based manually, and understanding the mutants and tool warnings completely would require deep understanding of all the software systems investigated.  However, in many cases, the impact of a mutant is clear, and the reason for warnings is obvious.  This was often enough the case that, as we discuss below, we are confident mutants that distinguish tools are meaningful (missed) opportunities for static analysis tools.
{\bf RQ6} and {\bf RQ7} receive only preliminary answers, for the smart contract tools.  We used the Universal Mutator tool for all experiments, since e.g., PIT \cite{pitmut} does not produce source-level mutants and the PMD tool requires source code; we also wanted to use a consistent mutation generation method for all languages, to the extent possible.

@@

For {\bf RQ4}, on the changes seen when restricting analysis to clean contracts, Slither did slightly better at detecting mutants when the original contract was clean for Slither, and the other two tools did somewhat worse on contracts for which they reported no findings.  For the three contracts clean for all tools, Slither performed almost exactly as it did over contracts in general, and the other tools performed worse, by about the same margin as they did for their own clean contracts.  For our approach, we only need a weak version of the ``clean program assumption'':  the threat is that kills may be under-reported for non-clean programs, due to interference with findings for the original code.  It is not a problem if mutation scores are \emph{worse} for programs where a tool reports no findings for the un-mutated code.  We therefore, for smart contracts, find no threat to our approach arising from the presence of findings on un-mutated code.  We speculate that ``clean'' results for some tools result from contracts where the tool has trouble with the contract code, but does not actually crash; Slither may do better on clean code because it has fewer such failures, and clean contracts are probably generally simpler and easier to analyze.

\subsubsection{Qualitative Examination of Mutants (RQ5)}

Because examining all mutants manually is infeasible, we focused primarily on examining mutants in the \emph{difference}.  Because our technique is differential, only mutants detected by at least one tool, but not detected by all tools, inform the comparison of static analysis tools.  The vast majority of the mutants in this set were meaningful.  We do not believe that all of them represent faults; some are harmless but unusual code changes.  A large number of the instances where use of {\tt tx.origin} in place of {\tt msg.sender} was flagged seem to us to be strange but not necessarily incorrect code.  On the other hand, at the risk of false positives, it is not unreasonable for tools to report some of these.   Our estimate is that, ignoring {\tt tx.origin} cases, at least 70\% of the mutants detected by one, but not all, tools, represent real bugs, and the failure to detect is roughly equally due to missing detectors and imprecise analysis.

\subsubsection{Improving Slither {\bf (RQ6 and RQ7)}}

@@

Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants, suggesting that for {\bf RQ7} it is indeed hard to use mutation analysis results without prioritization. A large majority of the mutants we inspected involved either the missing
{\tt return} problem noted in the introduction, or replacing {\tt msg.sender} with {\tt tx.origin}; Slither, of course, has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of the differences with respect to such changes are due to intentional behavior. SmartCheck and Securify tend to identify most (though not all) uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule, intended to reduce false positives.
It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ6} and {\bf RQ7}.

@@

\subsubsection{Qualitative Examination of Mutants (RQ5)}

For Java, again, the large majority of mutants we inspected where at least one tool detected the mutant and at least one tool did not detect it were definitely meaningful.  In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null} that would clearly introduce potential null pointer exceptions (NPEs).  While the differences between Solidity tools were often due to different detectors, the Java differences seemed mostly rooted in analysis engine methods; all tools aim to warn about potential NPEs.

@@

\subsubsection{Qualitative Examination of Mutants (RQ5)}

For Python, as with Java, many mutants we inspected involved deletion of a call, leading to a likely method call or access to a variable containing {\tt None}.  More Python mutants were more difficult to understand than for Solidity or Java, without a grasp of the larger program context, but a large number were still clearly differences in analysis engine capabilities, since all tools aim to detect failure to guard against improper uses of {\tt None}, the Python equivalent of a Java NPE.

@@

We used the Universal Mutator \cite{universalmutator,regexpMut}, which aggressively produces large numbers of mutants, but does not target any particular software defect patterns, to generate all mutants.  This tool has been heavily tuned and used for for Solidity and Python, and moderately tuned for Java.  The Java mutation also benefits from extensive tuning on C programs, which share syntactic mutation patterns with both Java and Solidity.
There is no room in the paper to present the exact set of projects analyzed, but we have provided an (anonymized) github repository containing raw results for inspection by reviewers, or further analysis by other researchers.  We will de-anonymize this if the paper is accepted.
